---
title: "Erik's Covid-19 Chart Pack"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

Data updated `r Sys.time()`.  World data are from [Worldometers](https://github.com/NovelCovid/API). National and state-level mortality, case, and testing data are from [Johns-Hopkins University](https://covidtracking.com). County and city-level mortality and case data are from the [New York Times](https://github.com/NovelCovid/API). Most data presented in this report were accessed through APIs provided by [The COVID Tracking Project](https://covidtracking.com) and [NovelCOVID API](https://github.com/NovelCovid/API).

```{r setup, include=FALSE}

# Dependencies
library(tidyverse)
library(dplyr)
library(ggplot2)
library(scales)
library(zoo)
#library(spData)
library(sf)
library(tmap)
library(kableExtra)
library(jsonlite)
library(urbnmapr)
library(geofacet)

# Global style settings
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
theme_set(theme_bw())
knitr::opts_chunk$set(fig.width=6, fig.height=4) 

# Import world data
continents <- fromJSON("https://disease.sh/v2/continents?yesterday=true") %>%
  arrange(continent)
countries <- fromJSON("https://disease.sh/v2/countries?yesterday=true") %>%
  mutate(iso_a3 = countryInfo$iso3)

# Import US data
usData <- read.csv("https://covidtracking.com/api/us/daily.csv")
usData$date = as.Date(as.character(usData$date), format = "%Y%m%d")
usData <- usData %>% 
  filter(date >= as.Date("2020-03-7")) %>%
  arrange(date) %>%
  mutate(activeCases = positive - recovered - death) %>%
  mutate(activeIncrease = activeCases - lag(activeCases,1)) %>%
  mutate(caseDeathRate = death/positive) %>%
  mutate(testPositiveRate = positiveIncrease/totalTestResultsIncrease) %>%
  mutate(dailyPositiveRate_ = positiveIncrease/totalTestResultsIncrease) %>%
  mutate(weekAverageDeaths = rollmean(deathIncrease,k=7,align="right", fill=NA)) %>%
  mutate(weekAverageCases = rollmean(positiveIncrease,k=7,align="right", fill=NA)) %>%
  mutate(weekAverageTests = rollmean(totalTestResultsIncrease,k=7,align="right", fill=NA)) %>%
  mutate(weekChangeDeaths = deathIncrease - lag(deathIncrease,7)) %>%
  mutate(weekChangeCases = positiveIncrease - lag(positiveIncrease,7)) %>%
  mutate(weekChangeTests = totalTestResultsIncrease - lag(totalTestResultsIncrease,7))
```

# World Data

There have been `r format(sum(countries$cases),big.mark = ",")` confirmed Covid-19 cases and `r format(sum(countries$deaths),big.mark = ",")` deaths worldwide.


```{r, fig.height=3.5, fig.width=6}
ggplot(continents, aes(x=continent,y=deaths)) +
  geom_col() +
  labs(title = "Deaths", y="Total Deaths") +
  scale_y_continuous(labels = comma) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.y = element_blank()) +
  coord_flip()
```

```{r, fig.height=3.5, fig.width=6}
ggplot(continents, aes(x=continent,y=cases)) +
  geom_col() +
  labs(title = "Cases", y="Total Cases") +
  scale_y_continuous(labels = comma) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.y = element_blank()) +
  coord_flip()
```

\newpage

```{r}
countryTable <- countries %>% 
  select(country,cases,deaths,todayCases,todayDeaths) %>%
  rename(Country = country, Cases = cases, Deaths = deaths, "New Cases" = todayCases, "New Deaths" = todayDeaths) %>%
  arrange(desc(Cases)) %>%
  head(20)

knitr::kable( 
  countryTable,
  caption = "Top Countries by Total Cases", 
  digits = 8, format = "latex",
  format.args = list(big.mark = ",")) %>% 
  kable_styling(latex_options = "hold_position")
```

\newpage

```{r, fig.width=6,fig.height=3.5}
data("World")
World$iso_a3 <- as.character(World$iso_a3)
World <- st_transform(World,4326)
worldMapData <- right_join(World,countries,by="iso_a3") 
worldMapData$deathsPerOneMillion[worldMapData$deathsPerOneMillion==0] = NA
worldMapData$casesPerOneMillion[worldMapData$casesPerOneMillion==0] = NA
worldMapData$testsPerOneMillion[worldMapData$testsPerOneMillion==0] = NA

tm_shape(worldMapData) + tm_borders() + 
  tm_fill(col = "deathsPerOneMillion", style = "log10", title = "Deaths per Million") +
  tm_layout(bg.color = "skyblue")
```

```{r, fig.width=6, fig.height=3.5}
tm_shape(worldMapData) + tm_borders() + 
  tm_fill(col = "casesPerOneMillion", style = "log10", title = "Cases per Million") +
  tm_layout(bg.color = "skyblue")
```

```{r, fig.width=6, fig.height=3.5}
tm_shape(worldMapData) + tm_borders() + 
  tm_fill(col = "testsPerOneMillion", style = "log10", title = "Tests per Million", palette = "BuGn") +
  tm_layout(bg.color = "skyblue")
```

\newpage

# National Data

There have been `r format(max(usData$positive),big.mark = ",")` confirmed Covid-19 cases and `r format(max(usData$death),big.mark = ",")` deaths in the United States.

```{r}
#usData %>% 
#  select(date,positive,death,positiveIncrease,deathIncrease) %>% 
#  arrange(desc(date)) %>%
#  top_n(n=7,wt=date) %>%
#  rename(Date = date, 
#         "Cases" = positive, 
#         "Deaths" = death, 
#         "New Cases" = positiveIncrease, 
#         "New Deaths" = deathIncrease)

counties <- get_urbn_map("counties", sf = TRUE) %>%
  st_transform(2163) %>%
  mutate(fips = as.integer(county_fips))

countyData <- fromJSON("https://disease.sh/v2/nyt/counties") %>%
  filter(date == max(date)) %>%
  mutate(fips = as.integer(fips))

countyDemographic <- read.csv("county_pop.csv") %>%
  select(STATE,COUNTY,STNAME,CTYNAME,POPESTIMATE2019) %>%
  mutate(fips = STATE*1000 + COUNTY)

countyMapData <- left_join(counties,countyData, by = "fips") 
countyMapData <- left_join(countyMapData,countyDemographic, by = "fips")
countyMapData <- countyMapData %>%
  mutate(deathsPerMil = 1e6*(deaths)/(POPESTIMATE2019)) %>%
  mutate(casesPerMil  = 1e6*(cases)/(POPESTIMATE2019))
countyMapData$deathsPerMil[countyMapData$deathsPerMil < 0.1] = 0.1
```


```{r}
usData %>% 
  select(date,positive,death,positiveIncrease,deathIncrease) %>% 
  arrange(desc(date)) %>%
  top_n(n=14,wt=date) %>%
  rename(Date = date, 
         "Cases" = positive, 
         "Deaths" = death, 
         "New Cases" = positiveIncrease, 
         "New Deaths" = deathIncrease) %>%
  knitr::kable(caption = "U.S. Deaths and Cases over the Last Two Weeks", digits = c(8,7,7,7,7,2,2), format = "latex", format.args = list(big.mark = ",")) %>% 
  kable_styling(latex_options = "hold_position")
```

\newpage

## Deaths

Because the effects of the virus can take several weeks to manifest in patients, deaths are a lagging indicator of contagion, but they may be a more reliable than case counts, which are a function of both the prevalence of the disease and the rate of testing. The case mortality rate is a very crude indicator of lethality because a large numbers of non-lethal cases are likely never detected. A declining case mortality rate is indicative of more widespread testing.

```{r}
ggplot(usData) + 
  geom_line(mapping = aes(x = date, y = death), size = 1) +
  labs(title = "Deaths", x = "Date", y = "Total Deaths") +
  scale_y_continuous(labels = comma, trans = "log10")
```

```{r}
ggplot(usData) +
  geom_line(mapping = aes(x = date, y = deathIncrease), alpha = 0.4) +
  geom_smooth(mapping = aes(x = date, y = deathIncrease), se=FALSE, span = 0.5) +
  labs(title = "New Deaths", x = "Date", y = "Deaths per Day") +
  scale_y_continuous(labels = comma)
```


```{r}
ggplot(usData, aes(x = date, y = deathIncrease/lag(positiveIncrease,14))) +
  geom_line(alpha = 0.5) +
  geom_smooth(se=FALSE, span = 0.5) +
  labs(title = "Daily Case Mortailty Rate", x = "Date", y = "Rate") +
  scale_y_continuous(labels = percent_format(), limits = c(0.0,0.25))
```

```{r}
ggplot(usData, aes(x = date, y = weekChangeDeaths)) + 
  geom_col(alpha = 0.5) +
  geom_smooth(se=FALSE, span = 0.5) +
  labs(title = "One-Week Change in Daily Deaths", x="Date", y="7-Day Change") +
  scale_y_continuous(labels = comma)
```

```{r, fig.width=7, fig.height=6}
tm_shape(countyMapData) + 
  tm_borders(lwd = 0.3) + 
#  tm_fill(col = "deathsPerTho",  style = "fixed", breaks = c(0,0.01,0.1,0.5,1.0,3.0), title = "Deaths Per 1K") +
  tm_fill(col = "deathsPerMil",  style = "log10", title = "Deaths Per 1M") +
  tm_layout(bg.color = "skyblue", legend.position = c("left","top"))
```

\newpage
## Cases

Reported cases are a function of both the spread of the disease and the prevalence of testing.

```{r}
ggplot(usData) +
  geom_line(mapping = aes(x = date, y = activeCases), size=1) +
  labs(title = "Active Cases", x = "Date", y = "Cases") +
  scale_y_continuous(labels = comma, trans = "log10")
```

```{r}
ggplot(usData) +
  geom_line(mapping = aes(x = date, y = positiveIncrease), alpha = 0.4) +
  geom_smooth(mapping = aes(x = date, y = positiveIncrease), se=FALSE, span = 0.5) +
  labs(title = "New Cases", x = "Date", y = "Cases per Day") +
  scale_y_continuous(labels = comma)
```

```{r}
ggplot(usData, aes(x = date, y = weekChangeCases)) + 
  geom_col(alpha = 0.5) +
  geom_smooth(se=FALSE, span = 0.5) +
  labs(title = "One-Week Change in Daily Cases", x="Date", y="7-Day Change") +
  scale_y_continuous(labels = comma)
```

```{r, fig.width=7, fig.height=6}
tm_shape(countyMapData) + 
  tm_borders(lwd = 0.3) + 
#  tm_fill(col = "casesPerTho", style = "fixed", breaks = c(0,0.1,1.0,5.0,10.0,30.0), title = "Cases Per 1K") +
  tm_fill(col = "casesPerMil", style = "log10", title = "Cases Per 1M") +
  tm_layout(bg.color = "skyblue", legend.position = c("left","top"))
```

\newpage
## Testing

Widespread testing is necessary for managing the spread of the disease. The following charts show how testing in the United States has changed over time. When the supply of available tests is limited, they are typically only used for patients whose symptoms suggest they are likely to have contracted the virus. A high positive test rate indicates that testing capacity is constrained.

```{r}
ggplot(usData) +
  geom_line(mapping = aes(x = date, y = totalTestResults),size=1) +
  labs(title = "Tests", x = "Date", y = "Total Number of Tests") +
  scale_y_continuous(labels = comma, trans = "log10")
```

```{r}
ggplot(usData) +
  geom_line(mapping = aes(x = date, y = totalTestResultsIncrease), alpha = 0.4) +
  geom_smooth(mapping = aes(x = date, y = totalTestResultsIncrease), se=FALSE, span=0.5) +
  labs(title = "New Tests", x = "Date", y = "Tests Per Day") +
  scale_y_continuous(labels = comma)
```


```{r}
ggplot(usData) +
  geom_line(mapping = aes(x = date, y = testPositiveRate), alpha = 0.4) +
  geom_smooth(mapping = aes(x = date, y = testPositiveRate), se=FALSE, span=0.5) +
  labs(title = "Positive Test Rate", x = "Date", y = "Rate per Day") +
  scale_y_continuous(labels = percent_format(2))
```

```{r}
ggplot(usData,  aes(x = date, y = weekChangeTests)) + 
  geom_col(alpha = 0.5) +
  geom_smooth(se=FALSE, span = 0.5) +
  labs(title = "One-Week Change in Daily Tests", x="Date", y="7-Day Change") +
  scale_y_continuous(labels = comma)
```

\newpage
# State Data


```{r}

# State name to FIPS code crosswalk
stateFIPS <- read.csv("state_FIPS_codes.csv") %>% 
  mutate(stusps = trimws(as.character(stusps)))

# Load state level historial data
stateData <- read.csv("https://covidtracking.com/api/v1/states/daily.csv") %>%
  mutate(state = as.factor(state)) %>%
  mutate(date = as.Date(as.character(date), format = "%Y%m%d"))

# Snampshot of most recent state data
currentStateData <- stateData %>% filter(date == max(date))

# State pupulatons
statePop = read.csv("state_pop.csv") %>%
  select(STATE,POPESTIMATE2019) %>%
  rename(st = STATE, population = POPESTIMATE2019)

# Create state data for mapping

# State map boundary and population data
states <- get_urbn_map("states", sf = TRUE) %>%
  mutate(st = as.integer(state_fips))
states <- left_join(states,statePop, by = "st")

#states <- us_states %>%
#  st_transform(2163) %>%
#  mutate(st = as.integer(GEOID)) 
# Merge current state covid data with geographic data via FIPS crosswalk
stateMapData <- currentStateData %>%
  select(state,positive,death,totalTestResults) %>%
  mutate(stusps = trimws(as.character(state)))
stateMapData <- full_join(stateMapData,stateFIPS, by="stusps")
stateMapData <- right_join(states,stateMapData,by="st")
# Per capita data for mappint
stateMapData <- stateMapData %>% 
  mutate(deathsPerMil = 1e6*death/population, 
         casesPerMil = 1e6*positive/population,
         testsPerMil = 1e6*totalTestResults/population,
         positiveTestRate = 100*positive/totalTestResults)

# Create list of 15 largest states by population
largeStates <- states %>%
  select(st, population) %>%
  arrange(desc(population)) %>%
  rename(fips = st) %>%
  head(15)

# Add variable to historical data for largest states and others
bigStateLvls = c(levels((stateData$state)),"Other")
stateData <- stateData %>%
  mutate(bigState = if_else(fips %in% largeStates$fips,
                            factor(state, levels = bigStateLvls),
                            factor("Other", levels = bigStateLvls)))

# Create summary data aggregating "Other" states together
bigStatData <- stateData %>%
  group_by(bigState,date) %>%
  summarize(death = sum(death, na.rm = TRUE),
          recovered = sum(recovered, na.rm = TRUE),
          deathIncrease = sum(deathIncrease, na.rm = TRUE),
          positive = sum(positive, na.rm = TRUE),
          positiveIncrease = sum(positiveIncrease, na.rm = TRUE),
          totalTestResults = sum(totalTestResults, na.rm = TRUE),
          totalTestResultsIncrease = sum(totalTestResultsIncrease, na.rm = TRUE))

# Add some derived variables
bigStatData <- bigStatData %>%
  arrange(bigState,date) %>%
  group_by(bigState) %>%
  mutate(activeCases = positive - recovered - death) %>%
  mutate(activeIncrease = activeCases - lag(activeCases,1)) %>%
  mutate(weekAverageDeaths = rollmean(deathIncrease,k=7,align="right", fill=NA)) %>%
  mutate(weekChangeDeaths = deathIncrease - lag(deathIncrease,7)) %>%
  mutate(weekAverageCases = rollmean(positiveIncrease,k=7,align="right", fill=NA))  %>%
  mutate(weekChangeCases = positiveIncrease - lag(positiveIncrease,7)) %>%
  mutate(weekAverageTests = rollmean(totalTestResultsIncrease,k=7,align="right", fill=NA))  %>%
  mutate(weekChangeTests = totalTestResultsIncrease - lag(totalTestResultsIncrease,7)) %>%
  mutate(positiveTestRate = positiveIncrease / totalTestResultsIncrease) %>%
  filter(date >= as.Date("2020-03-1"))
```

This section summarizes state-level data. Most data are reported for the largest 15 states by population, which account for `r format(100*sum(largeStates$total_pop_15)/sum(states$total_pop_15), digits = 1)` percent of the total U.S. population.

## Deaths

```{r}
topDeaths <- currentStateData %>%
  select(state,positive,death) %>%
  arrange(desc(death)) %>%
  top_n(n=10,wt=death)

topDeathsTable <- topDeaths %>%  
  rename( State = state,
          Cases = positive,
          Deaths = death)

```

```{r}
dateBreaks = c(as.Date("2020-3-1"), as.Date("2020-4-1"), as.Date("2020-5-1"), as.Date("2020-6-1"))

ggplot(bigStatData %>% filter(date == max(date)), aes(bigState,death)) +
  geom_col() +
  labs(title = "Deaths by State", y="Total Deaths") +
  scale_y_continuous(labels = comma) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.y = element_blank()) 
```

```{r, fig.height=6, fig.width=7}
ggplot(bigStatData, aes(x=date,y=weekAverageDeaths)) + 
  geom_line(alpha=0.4) +  
  geom_smooth(se=FALSE, span=0.5) +
  labs(title = "New Deaths by State", x = "Date", y = "Deaths per Day") +
  scale_y_continuous(labels = comma) +
  facet_wrap(~bigState, scales = "free_y") +
  scale_x_date(date_labels = "%b", breaks = dateBreaks)
```

```{r, fig.height=6, fig.width=7}
ggplot(bigStatData, aes(x=date,y=weekChangeDeaths)) + 
  geom_col(alpha=0.5) + 
  labs(title = "7-Day Change in Daily Deaths", x = "Date", y = "7-Day Change") +
  geom_smooth(se=FALSE, span=0.5) + 
  scale_y_continuous(labels = comma) +
  facet_wrap(~bigState, scales = "free_y") +
  scale_x_date(date_labels = "%b", breaks = dateBreaks)
```
\newpage

```{r, fig.height=5, fig.width=6}
tm_shape(stateMapData) + 
  tm_borders() + 
  tm_fill(col = "death", style = "log10", title = "Deaths") +    
  tm_layout(bg.color = "skyblue", legend.position = c("left","top"))
```

```{r, fig.height=5, fig.width=6}
tm_shape(stateMapData) + 
  tm_borders() + 
  tm_fill(col = "deathsPerMil", style = "log10", title = "Deaths per 1M") +
  tm_layout(bg.color = "skyblue", legend.position = c("left","top"))
```

\newpage
## Cases

```{r}
topCases <- currentStateData %>%
  select(state,positive,death) %>%
  arrange(desc(positive)) %>%
  top_n(n=10,wt=positive)

topCasesTable <- topCases %>%
  rename(State = state,
         Cases = positive,
         Deaths = death)
```


```{r}
ggplot(bigStatData %>% filter(date == max(date)), 
       aes(bigState,positive))  +
  geom_col() +
  labs(title = "Cases by State", y="Total Cases") +
  scale_y_continuous(labels = comma) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.y = element_blank())
```

```{r, fig.width=7, fig.height=6}
ggplot(bigStatData, aes(x=date,y=positiveIncrease)) +
  geom_line(alpha=0.4) +  
  geom_smooth(se=FALSE, span=0.5) + 
  labs(title = "New Cases by State", x = "Date", y = "Cases per Day") +
  scale_y_continuous(labels = comma) +
  facet_wrap(~bigState, scales = "free_y") +
  scale_x_date(date_labels = "%b", breaks = dateBreaks)
```

```{r, fig.width=7, fig.height=6}
ggplot(bigStatData, aes(x=date,y=weekChangeCases)) + 
  geom_col(alpha=0.5) +  
  labs(title = "7-Day Change in Daily Cases", x = "Date", y = "7-Day Change") +
  geom_smooth(se=FALSE, span=0.5) + 
  scale_y_continuous(labels = comma) +
  facet_wrap(~bigState, scales = "free_y") +
  scale_x_date(date_labels = "%b", breaks = dateBreaks)
```
\newpage

```{r, fig.height=5, fig.width=6}
tm_shape(stateMapData) + 
  tm_borders() + 
  tm_fill(col = "positive", style = "log10", title = "Cases") +
  tm_layout(bg.color = "skyblue", legend.position = c("left","top"))
```


```{r, fig.height=5, fig.width=6}
tm_shape(stateMapData) + 
  tm_borders() + 
  tm_fill(col = "casesPerMil", style = "log10", title = "Cases per 1M") +
  tm_layout(bg.color = "skyblue", legend.position = c("left","top"))
```

\newpage
## Testing

```{r}
ggplot(bigStatData %>% filter(date == max(date)), 
       aes(bigState,totalTestResults))  +
  geom_col() +
  labs(title = "Tests by State", y="Total Tests") +
  scale_y_continuous(labels = comma) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())
```


```{r, fig.width=7, fig.height=6}
ggplot(bigStatData, aes(x=date,y=totalTestResultsIncrease)) +
  geom_line(alpha=0.4) +  
  geom_smooth(se=FALSE, span=0.5) + 
  labs(title = "New Tests by State", x = "Date", y = "Tests per Day") +
  scale_y_continuous(labels = comma) +
  facet_wrap(~bigState, scales = "free_y") +
  scale_x_date(date_labels = "%b", breaks = dateBreaks)
```

\newpage
```{r, fig.width=6, fig.height=5}
tm_shape(stateMapData) + 
  tm_borders() + 
  tm_fill(col = "testsPerMil", style = "cont", title = "Tests per 1M", palette = "BuGn") +
  tm_layout(bg.color = "skyblue", legend.position = c("left","top"))
```

```{r, fig.width=6, fig.height=5}
tm_shape(stateMapData) + 
  tm_borders() + 
  tm_fill(col = "positiveTestRate", style = "cont", title = "Positive Rate", legend.format = list(digits=1)) +
  tm_layout(bg.color = "skyblue", legend.position = c("left","top"))
```
\newpage

Interpretation of differences in case rates across states is complicated by the fact that those states that do more thorough testing will invariably uncover more cases. A lower positive test rate is an indication that a state is doing more comprehensive testing since, when testing is rationed, only those individuals who are more likely to test positive are typically tested. The following chart compares the one-week increase in detected cases to the the number of tests administered by each state relative to population.  The states of greatest current concern are those with both a large increase in detected cases and a relatively small increase in tests.  These states lie in the upper-left of the chart.

```{r, fig.width=6.5, fig.height=6}
stateWatchData <- left_join(stateData, statePop %>% rename(fips=st), by = "fips")

stateWatchData <- stateWatchData %>%
  group_by(state) %>%
  arrange(state,date) %>%
  mutate(weekAverageDeaths = rollmean(deathIncrease,k=7,align="right", fill=NA)) %>%
  mutate(weekChangeDeaths = deathIncrease - lag(deathIncrease,7)) %>%
  mutate(weekAverageCases = rollmean(positiveIncrease,k=7,align="right", fill=NA))  %>%
  mutate(weekChangeCases = positiveIncrease - lag(positiveIncrease,7)) %>%
  mutate(weekAverageTests = rollmean(totalTestResultsIncrease,k=7,align="right", fill=NA))  %>%
  mutate(weekChangeTests = totalTestResultsIncrease - lag(totalTestResultsIncrease,7)) %>%
  mutate(positiveTestRate = positiveIncrease / totalTestResultsIncrease) %>%
  mutate(wkPositiveTestRate = (positive-lag(positive,7)) / (totalTestResults-lag(totalTestResults,7)) ) %>%
  mutate(wkCasesPerCap = 1e6*(positive-lag(positive,7)) / population) %>%
  mutate(wkTestsPerCap = 1e6*(totalTestResults-lag(totalTestResults,7)) / population) %>%
  filter(date >= max(date))
  
rates <- c(0.10,0.05,0.02)
ggplot(stateWatchData, aes(x=wkTestsPerCap,y=wkCasesPerCap, size=0.1*population,label=state)) + 
  geom_point(aes(x=wkTestsPerCap,y=wkCasesPerCap,size=population), alpha = 0.25) +
  geom_text() +
  geom_abline(intercept = 0, slope = rates[1], linetype = "dashed", size = 0.5, color = "gray") +
  geom_text(aes(15000,15000*rates[1]), label="10% Positive", size = 2.5, hjust = 1.1, color = "gray") +
  geom_abline(intercept = 0, slope = rates[2], linetype = "dashed", size = 0.5, color = "gray") +
  geom_text(aes(20000,20000*rates[2]), label="5% Positive", size = 2.5, hjust = 1.2, color = "gray") +
  geom_abline(intercept = 0, slope =rates[3], linetype = "dashed", size = 0.5, color = "gray") +
  geom_text(aes(20000,20000*rates[3]), label="2% Positive", size = 2.5, hjust = 1.3, color = "gray") +
  labs(title = "Tests vs. Cases by State", 
       x = "7-Day New Tests per Million", 
       y = "7-Day New Cases per Million",
       caption = "Point size proportional to state population.") +
  theme(legend.position = "none")
```

```{r}
ggsave("Tests_vs_Cases.png")
```


\newpage
# Local Data

The following charts and tables present mortality, case, and testing data for the Washington DC metropolitan area and adjacent states.


```{r}
localData <- stateData %>% 
  filter(state %in% c("VA","DC","MD")) %>%
  arrange(state,date) %>%
  group_by(state) %>%
  mutate(weekAverageDeaths = rollmean(deathIncrease,k=7,align="right", fill=NA)) %>%
  mutate(weekChangeDeaths = deathIncrease - lag(deathIncrease,7)) %>%
  mutate(weekAverageCases = rollmean(positiveIncrease,k=7,align="right", fill=NA))  %>%
  mutate(weekChangeCases = positiveIncrease - lag(positiveIncrease,7)) %>%
  mutate(positiveTestRate = ifelse(totalTestResultsIncrease > 100,
                                   positiveIncrease / totalTestResultsIncrease,
                                   NA)) %>%
  mutate(weekChangeTests = totalTestResultsIncrease - lag(totalTestResultsIncrease,7)) %>%
  filter(date >= as.Date("2020-03-1"))
```


```{r}
recentLocal <- localData %>%
  filter(date == max(date)) %>%
  select(state, positive, death, positiveIncrease, deathIncrease) %>%
  rename(State=state, "Cases" = positive, "Deaths" = death, "New Cases" = positiveIncrease, "New Deaths" = deathIncrease)
knitr::kable(recentLocal, caption = "Latest Local Data", format = "latex", format.args = list(big.mark = ",")) %>%
  kable_styling(latex_options = "hold_position")
```

\newpage
## Deaths

```{r}
ggplot(localData) + 
  geom_line(mapping = aes(x = date, y = death, color = state), size = 1) +
  labs(title = "Deaths", x = "Date", y = "Total Deaths", color = "State") +
  scale_y_continuous(labels = comma, trans = "log10") 
```

```{r, fig.width=6, fig.height=6}
ggplot(localData) +
  geom_line(mapping = aes(x = date, y = deathIncrease), alpha = 0.4) +
  geom_smooth(mapping = aes(x = date, y = deathIncrease), se=FALSE, span=0.5) +
  labs(title = "New Deaths", x = "Date", y = "Deaths per Day", color = "State") +
  facet_wrap(~state, ncol = 1, scales = "free_y") +
  scale_y_continuous(labels = comma)
```

```{r, fig.width=6, fig.height=6}
ggplot(localData, aes(x = date, y = weekChangeDeaths)) + 
  geom_col(alpha = 0.5) +
  geom_smooth(se=FALSE, span=0.5) + 
  labs(title = "One-Week Change in Daily Deaths", x="Date", y="7-Day Change") +
  facet_wrap(~state, ncol = 1, scales = "free_y") +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "none")
```

```{r, fig.width=6, fig.height=5}
localMapData <- countyMapData %>%
  filter(state_fips %in% c(24,11,51)) 
localStateMaps <- stateMapData %>%
  filter(st %in% c(24,11,51))

tm_shape(localMapData) + 
  tm_borders(lwd = 0.3) + 
  tm_fill(col = "deathsPerMil", style = "log10", title = "Deaths per 1M") +
  tm_shape(localStateMaps) +
  tm_borders(lwd = 1.5, col = "gray") +
  tm_layout(bg.color = "skyblue")
```

```{r, fig.width=6, fig.height=5}
nearMapData <- localMapData %>%
  filter(fips %in% c(51600,51510,51610,51059,51107,51013,11001,24031,24033))

tm_shape(nearMapData) + 
  tm_borders() + 
  tm_fill(col = "deathsPerMil", style = "log10", title = "Deaths per 1M") +
  tm_text("county", size=0.5) +
  tm_layout(bg.color = "skyblue")
```

\newpage
## Cases

```{r}
ggplot(localData) + 
  geom_line(mapping = aes(x = date, y = positive, color = state), size=1) +
  labs(title = "Cases", x = "Date", y = "Total Cases", color = "State") +
  scale_y_continuous(labels = comma, trans = "log10") 
```

```{r, fig.width=6, fig.height=6}
ggplot(localData) +
  geom_line(mapping = aes(x = date, y = positiveIncrease), alpha=0.4) +
  geom_smooth(mapping = aes(x = date, y = positiveIncrease), se=FALSE, span=0.5) +
  labs(title = "New Cases", x = "Date", y = "Cases per Day", color = "State") +
  facet_wrap(~state, ncol = 1, scales = "free_y") +
  scale_y_continuous(labels = comma)
```

```{r, fig.width=6, fig.height=6}
ggplot(localData, aes(x = date, y = weekChangeCases)) + 
  geom_col(alpha = 0.5) +
  geom_smooth(se = FALSE, span = 0.5) +
  labs(title = "One-Week Change in Daily Cases", x="Date", y="7-Day Change") +
  facet_wrap(~state, ncol = 1, scales = "free_y") +
  scale_y_continuous(labels = comma) +
  theme(legend.position = "none")
```

```{r, fig.width=6, fig.height=5}
tm_shape(localMapData) + 
  tm_borders(lwd = 0.3) + 
  tm_fill(col = "casesPerMil", style = "log10", title = "Cases per 1M") +
  tm_shape(localStateMaps) +
  tm_borders(lwd = 1.5, col = "gray") +
  tm_layout(bg.color = "skyblue")
```

```{r, fig.width=6, fig.height=5}
tm_shape(nearMapData) + 
  tm_borders() + 
  tm_fill(col = "casesPerMil", style = "log10", title = "Cases per 1M") +
  tm_text("county", size=0.5) +
  tm_layout(bg.color = "skyblue")
```

\newpage
## Testing

```{r}
ggplot(localData) + 
  geom_line(mapping = aes(x = date, y = totalTestResults, color = state), size = 1) +
  labs(title = "Tests", x = "Date", y = "Total Tests", color = "State") +
  scale_y_continuous(labels = comma, trans = "log10") 
```

```{r, fig.width=6, fig.height=6}
ggplot(localData, aes(x = date, y = totalTestResultsIncrease)) +
  geom_line(alpha=0.4) +
  geom_smooth(se=FALSE, span=0.5) +
  labs(title = "New Tests", x = "Date", y = "Tests per Day") +
  facet_wrap(~state, scales = "free_y", ncol = 1) + 
  scale_y_continuous(labels = comma)
```

```{r, fig.width=6, fig.height=6}
ggplot(localData, aes(x = date, y = positiveTestRate)) +
  geom_line(alpha=0.4) +
  geom_smooth(se=FALSE, span=0.5) +
  labs(title = "Positive Test Rate", x = "Date", y = "Rate") +
  scale_y_continuous(labels = percent_format(2), limits = c(0.0,0.4)) +
  facet_wrap(~state, ncol = 1) 
```
